### 2017-02-02 manuscript.rmd
plot formatting, increased dup_adjust to 5

	# These clade dates are from the KMRR code, they got them from http://timetree.org/
	calibration_times = data.frame(
		age.min =
			c(  7, 9, 29, 92, 162, 167, 296 ), 
		clade = 
			c( "Hominini", "Homininae", "Catarrhini", "Euarchontoglires", "Theria", "Mammalia", "Amniota" ),
		stringsAsFactors=FALSE
	)
```

@@

	calibration_times_all = data.frame(
        age =
            c( 20, 92, 167, 9, 42, 74, 296, 535, 104, 937, 29, 722, 441, 65, 15, 1215, 
                414, 371, 162, 25, 74, 77, 86, 7 ), 
        clade = 
            c("Hominoidea", "Euarchontoglires", "Mammalia", "Homininae", 
                    "Simiiformes", "Primates", "Amniota", "Vertebrata", "Eutheria", 
                    "Bilateria", "Catarrhini", "Chordata", "Euteleostomi", 
                    "Haplorrhini", "Hominidae", "Opisthokonta", "Sarcopterygii", 
                    "Tetrapoda", "Theria", "Murinae", "Sciurognathi", "Rodentia", "Glires", "Hominini"
            ),
        stringsAsFactors=FALSE
    )

	pairwise_summary %>% 
		filter( species_a == "homo_sapiens" | species_b == "homo_sapiens" ) %>% 
		filter( ! is.na( D ) ) %>%
		group_by( Event, label ) %>% 
		summarise( cor.test(  tau_a, tau_b )$estimate, n=n() ) %>% 
		rename( rho=`cor.test(tau_a, tau_b)$estimate` ) %>%
		rename( clade = label ) %>% 
		left_join( calibration_times_all, clade=clade ) %>%
		ggplot() + 
			geom_point( aes( x=age, y=rho, col=Event, size=n ) ) +
			theme_classic()

@@

In order to better understand why our phylogenetic analysis supports a different conclusion (*i.e.*, no support for the ortholog conjecture) than the published analysis of KMRR [@KryuchkovaMostacci:2016iw] (*i.e.*, strong support for the ortholog conjecture), we first checked to make sure we could reproduce their result based on pairwise analyses. This is important since we are only looking at at a subset of the data they considered, the Brawand *et al.* 2011 [@Brawand:2011du] dataset for gene trees that could be successfully time calibrated. We can replicate their result - we find a higher Tau correlation between orthologs (R=`r round( ortholog_r, 3) `) than between paralogs (R=`r round( paralog_r, 3)`).

@@

The reason that paralog pairs have Tau values that are more different than ortholog pairs is that they tend to be more distantly related, not because there are different rates of Tau evolution following duplication and speciation events. This means that having information on whether two genes are orthologs or paralogs provides little added information about expression beyond knowing how distantly related the two genes are. This has several implications. It is an example of the limitations of the concepts of orthology and paralogy [@Dunn:2016er]. These terms can have straightforward meaning in small gene trees with simple duplication/speciation histories, but the utility of the terms breaks down on larger more complex gene trees. Orthology and paralogy are annotations on the tips of the tree that are derived from the inferred structure of the tree and history of duplication and speciation at internal tree nodes. In this sense, orthology and paralogy are statements about the internals of the tree that are distilled into statements at the tips of the tree. Much is lost in the process, though. These are the same pre-computed trees we use in our phylogenetic re-analyses of their study. For most questions it is much more direct to focus on the structure of the tree and the inferred processes within the tree, such as which internal nodes are duplication or speciation events and how much change occurs along the branches. In this case, KMRR [@KryuchkovaMostacci:2016iw] relied on orthology and paralogy annotations that are derived from the inferred duplication and speciation history on gene phylogenies [@Herrero:2016jj].

@@

	# Perform contrasts 
	gene_trees_sim_null %<>% add_pics_to_trees()
	nodes_sim_null_contrasts = gene_trees_sim_null %>% summarize_contrasts()

@@

	dup_adjust = 2

@@

	# Calculate contrasts 
	gene_trees_sim_oc %<>% add_pics_to_trees()
	nodes_sim_oc_contrasts = gene_trees_sim_oc %>% summarize_contrasts()

@@

	g_phylo_null = 
		nodes_sim_null_contrasts %>% 
		ggplot( aes( x=pic, y=..density.., col=Event ) ) + 
			geom_freqpoly( binwidth=.001, position="identity" ) +
			xlab( "Phylogenetic independent \ncontrasts (PIC)" ) +
			ylab( "Density" ) +
			xlim( 0, 0.075 ) + 
			ylim( 0, 100 ) +
			theme_classic() + 
			ggtitle( "(b) Phylogenetic result under null simulation" )

	g_pair_null = 
		pairwise_sim_null_summary %>% 
		filter( species_a == "homo_sapiens" | species_b == "homo_sapiens" ) %>% 
		filter( ! is.na( D ) ) %>%
		group_by( Event, label ) %>% 
		summarise( cor.test(  tau_a, tau_b )$estimate, n=n() ) %>% 
		rename( rho=`cor.test(tau_a, tau_b)$estimate` ) %>%
		rename( clade = label ) %>% 
		left_join( calibration_times_all, clade=clade ) %>%
		ggplot() + 
			geom_point( aes( x=age, y=rho, col=Event, size=n ) ) +
			theme_classic() +
			ggtitle( "(a) Pairwise result under null simulation" )

	g_phylo_oc = 
	nodes_sim_oc_contrasts %>% 
	ggplot( aes( x=pic, y=..density.., col=Event ) ) + 
		geom_freqpoly( binwidth=.001, position="identity" ) +
		xlab( "Phylogenetic independent \ncontrasts (PIC)" ) +
		ylab( "Density" ) +
		xlim( 0, 0.075 ) + 
		ylim( 0, 100 ) +
		theme_classic() + 
		ggtitle( "(d) Phylogenetic result under OC simulation" )

	g_pair_oc = 
		pairwise_sim_oc_summary %>% 
		filter( species_a == "homo_sapiens" | species_b == "homo_sapiens" ) %>% 
		filter( ! is.na( D ) ) %>%
		group_by( Event, label ) %>% 
		summarise( cor.test(  tau_a, tau_b )$estimate, n=n() ) %>% 
		rename( rho=`cor.test(tau_a, tau_b)$estimate` ) %>%
		rename( clade = label ) %>% 
		left_join( calibration_times_all, clade=clade ) %>%
		ggplot() + 
			geom_point( aes( x=age, y=rho, col=Event, size=n ) ) +
			theme_classic() +
			ggtitle( "(c) Pairwise result under OC simulation" )

@@

We simulated a data set under the null model that tau evolves at the same rate following duplication and speciation events. Under the null model, pairwise analysis results (Figure XXSim a) are very similar as for the observed data, with higher rho for orthologs (`r ortholog_r_sim_null`) than paralogs (`r paralog_r_sim_null`). In the phylogenetic analysis of the null simulation, contrasts show similar divergence following speciation and duplication (Figure XXSim b) and does not reject the null hypothesis (Wilcoxon p=`r wilcox_oc( nodes_sim_null_contrasts )`).

We next simulated a data set under the ortholog conjecture, where the rate of evolution of Tau following duplication was `r dup_adjust` fold the rate following speciation. The results of this heterogeneous model are nearly indistinguishable from the null model  (Figure XXSim c), and also have a higher rho for orthologs (`r ortholog_r_sim_oc`) than paralogs (`r paralog_r_sim_oc`). In the phylogenetic analysis of the ortholog conjecture simulation, contrasts XX (Figure XXSim c) and XX (Wilcoxon p=`r wilcox_oc( nodes_sim_oc_contrasts )`).

@@

>>>>

	# These clade dates are from the KMRR code, they got them from http://timetree.org/
	# only the following are used in phylogenetic analyses - 
	# c( "Hominini", "Homininae", "Catarrhini", "Euarchontoglires", "Theria", "Mammalia", "Amniota" )
	calibration_times = data.frame(
        age =
            c( 20, 92, 167, 9, 42, 74, 296, 535, 104, 937, 29, 722, 441, 65, 15, 1215, 
                414, 371, 162, 25, 74, 77, 86, 7 ), 
        clade = 
            c("Hominoidea", "Euarchontoglires", "Mammalia", "Homininae", 
                    "Simiiformes", "Primates", "Amniota", "Vertebrata", "Eutheria", 
                    "Bilateria", "Catarrhini", "Chordata", "Euteleostomi", 
                    "Haplorrhini", "Hominidae", "Opisthokonta", "Sarcopterygii", 
                    "Tetrapoda", "Theria", "Murinae", "Sciurognathi", "Rodentia", "Glires", "Hominini"
            ),
        stringsAsFactors=FALSE
    )

	# In simulation, the fold change in rate following duplication relative to speciation
	dup_adjust = 5

@@

	pairwise_summary %>% 
		filter( species_a == "homo_sapiens" | species_b == "homo_sapiens" ) %>% 
		filter( ! is.na( D ) ) %>%
		group_by( Event, label ) %>% 
		summarise( cor.test(  tau_a, tau_b )$estimate, n=n() ) %>% 
		rename( rho=`cor.test(tau_a, tau_b)$estimate` ) %>%
		rename( clade = label ) %>% 
		left_join( calibration_times, clade=clade ) %>%
		ggplot() + 
			geom_point( aes( x=age, y=rho, col=Event, size=n ) ) +
			theme_classic()

@@

In order to better understand why our phylogenetic analysis supports a different conclusion (*i.e.*, no support for the ortholog conjecture) than the published analysis of KMRR [@KryuchkovaMostacci:2016iw] (*i.e.*, strong support for the ortholog conjecture), we first checked to make sure we could reproduce their result based on pairwise analyses. This is important since we are only looking at at a subset of the data they considered, the Brawand *et al.* 2011 [@Brawand:2011du] dataset for gene trees that could be successfully time calibrated. In their Figure 1, they present a higher Tau correlation coefficient between ortholog pairs than paralog ortholog pairs. We find the same here, with correlation coefficients of R=`r round( ortholog_r, 3)` between ortholog pairs and `r round( paralog_r, 3)` for paralog pairs.

@@

@@

	# Perform contrasts 
	gene_trees_sim_null %<>% add_pics_to_trees()
	nodes_sim_null_contrast = gene_trees_sim_null %>% summarize_contrasts()

@@

@@

	# Calculate contrasts 
	gene_trees_sim_oc %<>% add_pics_to_trees()
	nodes_sim_oc_contrast = gene_trees_sim_oc %>% summarize_contrasts()

@@

	g_phylo_null = 
		nodes_sim_null_contrast %>% 
		ggplot( aes( x=pic, y=..density.., col=Event ) ) + 
			geom_freqpoly( binwidth=.001, position="identity" ) +
			xlab( "Phylogenetic Independent Contrast" ) +
			ylab( "Density" ) +
			xlim( 0, 0.05 ) + 
			ylim( 0, 75 ) +
			theme_classic() + 
			ggtitle( "(b) Null simulation, phylogenetic" ) +
			theme(legend.justification=c(1,1), legend.position=c(1,1))

	g_pair_null = 
		pairwise_sim_null_summary %>% 
		filter( species_a == "homo_sapiens" | species_b == "homo_sapiens" ) %>% 
		filter( ! is.na( D ) ) %>%
		group_by( Event, label ) %>% 
		summarise( cor.test(  tau_a, tau_b )$estimate, n=n() ) %>% 
		rename( rho=`cor.test(tau_a, tau_b)$estimate` ) %>%
		rename( clade = label ) %>% 
		left_join( calibration_times, clade=clade ) %>%
		ggplot() + 
			geom_point( aes( x=age, y=rho, col=Event, size=n ) ) +
			theme_classic() +
			xlab( "Age" ) +
			ylab( "Rho" ) +
			ggtitle( "(a) Null simulation, pairwise" ) +
			theme(legend.position="none")

	g_phylo_oc = 
	nodes_sim_oc_contrast %>% 
	ggplot( aes( x=pic, y=..density.., col=Event ) ) + 
		geom_freqpoly( binwidth=.001, position="identity" ) +
		xlab( "Phylogenetic Independent Contrast" ) +
		ylab( "Density" ) +
		xlim( 0, 0.05 ) + 
		ylim( 0, 75 ) +
		theme_classic() + 
		ggtitle( "(d) OC simulation, phylogenetic" ) +
		theme(legend.position="none")

	g_pair_oc = 
		pairwise_sim_oc_summary %>% 
		filter( species_a == "homo_sapiens" | species_b == "homo_sapiens" ) %>% 
		filter( ! is.na( D ) ) %>%
		group_by( Event, label ) %>% 
		summarise( cor.test(  tau_a, tau_b )$estimate, n=n() ) %>% 
		rename( rho=`cor.test(tau_a, tau_b)$estimate` ) %>%
		rename( clade = label ) %>% 
		left_join( calibration_times, clade=clade ) %>%
		ggplot() + 
			geom_point( aes( x=age, y=rho, col=Event, size=n ) ) +
			theme_classic() +
			xlab( "Age" ) +
			ylab( "Rho" ) +
			ggtitle( "(c) OC simulation, pairwise" ) +
			theme(legend.position="none")

@@

We simulated a data set under the null model that tau evolves at the same rate following duplication and speciation events. Under the null model, pairwise analysis results (Figure XXSim a) are very similar as for the observed data, with higher rho for orthologs (`r ortholog_r_sim_null`) than paralogs (`r paralog_r_sim_null`). In the phylogenetic analysis of the null simulation, contrasts show similar divergence following speciation and duplication (Figure XXSim b) and does not reject the null hypothesis (Wilcoxon p=`r wilcox_oc( nodes_sim_null_contrast )`).

We next simulated a data set under the ortholog conjecture, where the rate of evolution of Tau following duplication was `r dup_adjust` fold the rate following speciation. The results of this heterogeneous model are nearly indistinguishable from the null model  (Figure XXSim c), and also have a higher rho for orthologs (`r ortholog_r_sim_oc`) than paralogs (`r paralog_r_sim_oc`). In the phylogenetic analysis of the ortholog conjecture simulation, contrasts XX (Figure XXSim c) and XX (Wilcoxon p=`r wilcox_oc( nodes_sim_oc_contrast )`).

@@

We find that having information on whether two genes are orthologs or paralogs provides little added information about expression beyond knowing how distantly related the two genes are. This has several implications. It is an example of the limitations of the concepts of orthology and paralogy [@Dunn:2016er]. These terms can have straightforward meaning in small gene trees with simple duplication/speciation histories, but the utility of the terms breaks down on larger more complex gene trees. Orthology and paralogy are annotations on the tips of the tree that are derived from the inferred structure of the tree and history of duplication and speciation at internal tree nodes. In this sense, orthology and paralogy are statements about the internals of the tree that are distilled into statements at the tips of the tree. Much is lost in the process, though. These are the same pre-computed trees we use in our phylogenetic re-analyses of their study. For most questions it is much more direct to focus on the structure of the tree and the inferred processes within the tree, such as which internal nodes are duplication or speciation events and how much change occurs along the branches. In this case, KMRR [@KryuchkovaMostacci:2016iw] relied on orthology and paralogy annotations that are derived from the inferred duplication and speciation history on gene phylogenies [@Herrero:2016jj].
